---
title: |
  | An Application of Bayesian Statistical Analysis:
  | Effects of Party Brand on Party Identification
author:
- affiliation: Department of Political Science, University of Illinois at Urbana-Champaign.
    <sk12@illinois.edu>
  name: Sanghoon Kim
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    keep_tex: yes
    md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
    number_sections: yes
    template: C:/Users/saoon/Documents/Quant 2 2017/531-explorations/bowersarticletemplate.latex
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 7
    md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
fontfamily: Crimson Text
fontsize: 12pt
geometry: margin=1in
header-includes: \usepackage{setspace}\doublespacing
bibliography: finalpaper.bib
biblio-style: apsr

---

```{r include=FALSE, cache=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.

## To make the html file do
## render("exploration4.Rmd",output_format=html_document(fig_retina=FALSE))
## To make the pdf file do
## render("exploration4.Rmd",output_format=pdf_document())

require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",    # slightly smaller font for code
  echo=FALSE,
  results='markup',
  strip.white=TRUE,
  fig.path='figs/fig',
  cache=FALSE,
  highlight=TRUE,
  width.cutoff=132,
  size='footnotesize',
  out.width='.9\\textwidth',
  fig.retina=FALSE,
  message=FALSE,
  warning=FALSE,
  comment=NA)
```

```{r}
load("replication.rda")
```

# Introduction

"Questions of causality rather than questions of description or association" have dominated recent researches and guided researchers into better understanding on the causal inference from statistical analyses [@imbens2015, xvii]. Experimental designs have provided powerful tools for answering those questions in their quests for causal inferences with increasing sophistication in research design and technological improvement. With the development of experimental methods based on treatment and control conditions, scholars have also focused on the potential outcomes of the experiments to better understand the concept of causality. Because it is impossible to observe both treatment and control outcomes from a single subject, scholars have devised different statistical sohpistications to measure treatment effects of experiments. One of those experimental designs is used in Lupu (2013) article, "Party brands and partisanship: Theory with evidence from a survey experiment in Argentina," providing average treatment effects as a result of an experiment in Argentina. The article approaches to the question of party ID with a Bayesian psychological theory that people with low information will be affected by a treatment more than those with high information. This term paper develops Lupu's article by conducting a Bayesian statistical analysis to examine their empirical findings. We approach to this analysis by treating potential outcomes as random variables and applying model-based inference to an experimental design [@imbens2015]. We find contradicting results with the Bayesian analysis. Contrary to the prediction and findings that low information prior results in high treatment effects, we find the effects are not highly related to the information prior. 


# Overview of the Original Article
\setlength{\parindent}{0.7cm}
How will changes in party identity affect party attachments of voters? Scholars have approached to this question with various theoretical perspectives and distinctive empirical findings. Lupu (2013) approaches to the question with an experimental design of party brand, emphasizing the importance in the supply side of the party attachment. The article develops a model of party brands and examines how a political party's behavior affects citizen’s attachments to the party [@Lupu2013]. Partisan identification is expected to diminish when a party of support converges with other parties and becomes indiscriminable from other parties, while it increases when different parties diverge each other in their ideological stances. They find empirical evidence for the theory from a survey experiment, which is fielded in Cordoba and Santa Fe in Argentina. The experiment tests three treatments: a treatment of comparing each party’s platform on important issues, another of providing information on interparty alliance and party switching, and the last one showing both first and second treatments. The first treatment expects to see an increase in partisan identification, while the second treatment predicts a decrease. They find the expected results from the experiment, a positive change with the first and a negative change with the second treatment. They also find evidences for a Bayesian psychological theory that voters with low information on parties are more likely to be affected by treatments, vice versa. Younger voters and people with low information show higher treatment effects from the experiment.

The research presents straightforward results from an experimental design. They find the average treatment effect by subtracting the proportion of people identified with any party in a treatment group with those in a control group. In addition, their theoretical prediction is verified by estimating conditional average treatment effects (CATEs), results of which support the Bayesian psychological prediction with the information prior. 

While CATEs present evidences for the main theoretical argument, they are limited in providing a more thorough empirical support for Bayesian analysis. The assumptions behind CATEs are based on the prior and posterior information of voters, but this article does not advance in implementing actual Bayesian statistical analysis. This lead us to expect that conducting actual Bayesian analysis will provide us a better interpretation and understanding of the experiment. Thus, this paper aims to enrich Lupu (2013)’s statistical analysis by adopting a Bayesian statistical approach with  following the literature on causal inference [@imbens2015; @gelman2014]. In an experimental research design, scholars have suggested different methods accounting for counterfactual observations between treatment and control. Once a target receives a treatment, it is impossible to observe a control response from the same person. Under Fisher's and Neyman’s methods, different approaches have been applied to fill in the 'blanks' of potential outcomes. Apart from those two methods, Bayesian analysis introduces a model-based approach, which estimates the distribution of missing data from observed data and parameters of interest. While this paper's application of Bayesian methods may be limited in its details, we attempt to follow this method and implement statistical inference for the original paper. We use age treatment assignment as the prior distribution and derive outcomes based on linear model and logistic model approaches. Results from Bayesian analysis show differing outcomes from the original paper, which calls for additional attention. 

This paper proceeds as follows. We first analyze the operating characteristics of the procedures, such as bias, Type I error and Power test, and confidence intervals. Bayesian statistical approaches follows with a more detailed description of the motivation behind the analysis taken in this paper. This paper concludes with some final remarks. 



# Analysis of the operating characteristics of the procedures

This section presents analysis of the operating characteristic of statistical inference of the original paper. Lupu (2013) basically utilizes a difference-in-means method, subtracting differences in proportions of partisan identification before and after treatments. The empirical analysis also has three categories of treatment, two main treatments and one combined treatment. We focus on the first main treatment of policy platform and check the bias, Type I error, power of the tests, and confidence intervals based on an ordinary least squares method of main variables. 

## Bias Test
An OLS analysis results the smallest bias in its design and running a bias test seems redundant for statistical inference. The following results confirm the assumption about biasness regarding the OLS approach showing minimal bias with the linear model. The density plot in Figure \ref{fig:bias} indicates unbiasedness of the main OLS process and the qqplot also confirms this finding. 

\begin{figure}[H]
\centering
\includegraphics[height=6cm, width=12cm]{bias graph.pdf}
\caption{Bias Test Graph}\label{fig:bias}
\end{figure}

## Type I Error 
Type I error refers to the rates that we reject falsely correct hypothesis. To estimate this rate, we disconnect the relationship between outcome and explanatory variables by shuffling the treatment variable. Then we get a coefficient from each test and examine the proportion of p-values smaller than 0.05. If this rate is larger than 0.05, we are rejecting the hypothesis more than it is supposed to and we can say we have high Type I error. We expect to see this rate around 0.05. The result corresponds to this criterion and returns the value of 0.052. 

## Power of test
To estimate the power of tests, we start from subtracting treatment effects from the outcome variable and make a world of known effects by shuffling the treatment variable. From the shuffled data, we get a distribution p-values and check the proportion of p-values larger than 0.05. The end result returns the power of tests. Different values from the original linear model, a robust linear model, and a binomial model are presented. Results indicate that both linear and robust linear models have power of around 0.95, showing low Type II error rates. In contrast, the power is low with the logistic analysis with a value of 0.16. 

## Confidence Interval 
For confidence intervals, this paper tests two different methods: a CLT-IID based method and a randomization-based method. For the former method, we derive confidence intervals from p-values of the canned linear regression results. For the latter one, we use inverting tests to make confidence intervals, based on p-values from Fisher’s methods of comparing coefficients from permutation and the observed coefficient. In the inverting test method, we find confidence intervals by finding out points where p-values are larger than 0.05. These points create a region that we cannot reject the null hypothesis of no effect, which equals to the confidence intervals of interests. For the randomization-based test, this method returns confidence intervals with better coverage. 

The confidence intervals from canned regression results are [0.008843242 0.166580419] and the coefficient from our baseline regression (0.0877) falls within this region. In contrast, confidence intervals from the second method diverges from the previous intervals, ranging from -0.16762768 to -0.01172034. Figure 2 of Lupu (2013) article provides the average treatment effects with confidence intervals, with the x-axis indicating percentage treatment effects. The original confidence intervals match the CLT-IID based ones but not results from the randomization-based method. The difference between these two methods may indicate the non-normality of coefficient distribution. This can be found with the following graph (Figure \ref{fig:density}) from the CLT based method: the distribution is skewed to the right and clustered on the left side. In short, the divergence in confidence intervals requires researchers to take an additional attention to approach to this experiment data.

\begin{figure}[H]
\centering
\includegraphics[width=.75\textwidth]{nonnormal.pdf}
\caption{CLT-IID based Distribution of Coefficients}\label{fig:density}
\end{figure}

We also derive a p-value from the permutation-based approach. The outcome is significant at 0.05 level with the p-value of 0.032. A final remark remains with the consistency test. We implemented consistency tests with different methods but we found limited results from them. The code appendix provides the process we followed to test consistency. 


# Implementing Bayesian Statistical Analysis
Imbens and Rubin implements Bayesian model-based imputation to fill in the missing observations given observed outcome variable and treatment [@imbens2015]. They generate the distribution of the missing data given the observed data based on the understanding that potential outcomes are also random variables. Therefore, an anlaysis on the distribution of the missing variable centers on their statistical analysis. Table \ref{table:age} shows the proportion of treatment for each age group. Lupu (2012) finds conditional average treatment effect of age by simply dividing the entire respondents by age groups and calculated treatment effects for each group. Results confirmed theoretical model of party brands from the original article. Table \ref{table:age} presents a slight disparity in the number of subjects in each group as well as the proportion of treatment and control. The rate of people received treatment tends to decrease as we move to older age groups. 

\begin{table}[]
\centering
\caption{Age and Treatment Proportions}
\label{table:age}
\begin{tabular}{l|lllll} \hline
      & \textless=20 & 30-40 & 40-50 & 50-60 & \textgreater=60 \\ \hline
0     & 63           & 62    & 55    & 47    & 71              \\
1     & 66           & 69    & 53    & 44    & 59              \\
Total & 129          & 131   & 108   & 91    & 130            \\
Treatment Proportion & 0.51 & 0.53 & 0.49 & 0.48 & 0.45 \\ \hline \hline
\end{tabular}
\end{table}



\begin{table}[]
\centering
\caption{Partisanship Among Age Groups}
\label{table:partisan}
\begin{tabular}{l|lllll} \hline
      & \textless=20 & 30-40 & 40-50 & 50-60 & \textgreater=60 \\ \hline
0     & 45           & 40    & 32    & 31    & 40              \\
1     & 15           & 22    & 23    & 16    & 31              \\
Partisanship Proportion & 0.24 & 0.35 & 0.42 & 0.34 & 0.44 \\ \hline \hline
\end{tabular}
\end{table}

Table \ref{table:partisan} presents an additional concern. The proportion of partisans is disparate among different age groups and younger citizens are more disinterested in politics than older generations. Therefore, changes in party identification from the experiment may signify different interpretations for each group. It is possible that because of disparity in prior distribution of partisanship across age groups, the suggested treatment effects may not represent different information priors but just show inherent characteristics of generational differences. In short, the original research fails to account for the disparity in experiment assignments and the distribution of outcome variable across age groups. 

We hope to find evidences for the suggested criticism with a Bayesian statistical approach. We conduct three analyses of the approach: first, we conduct a Bayesian analysis with normal prior and posterior distribution.We present the posterior distribution by using the Rstan package in R. Second, we carry out a multilevel Bayesian analysis with the normal distribution following two specific methods. We first use the original treatment variable as our explanatory variable and, second, we move on to use the global distance of the original treatment variable as our new explanatory variable. For the multiple levels, the age groups provide the 'local' level information about the different treatment proportions and then the actual treatment effects are analyzed at the 'population' level. Third, we examine another multilevel Bayesian anlysis based on the bernoulli-binomial distribution. 


## Basic Bayesian Analysis

Figure \ref{fig:basic} shows the posterior distribution from a single-level Bayesian analysis. The left panel indicates the distribution of intercept and the right one shows the coefficient for the first treatment. Results from this first analysis matches results from the original article. The mean coefficient from the posterior distribution is 0.0879 and the intercept is 0.3580, which are comparable to 0.0877 and 0.3579 from the OLS results. Figure \ref{fig:basic} also shows mode values of intercept and treatment coefficient (0.3481 and 0.0863, respectively). 

\begin{figure}[H]
\centering
\includegraphics[height=6cm, width=12cm]{figunnamed-chunk-4-1.pdf}
\caption{Basic Bayesian Analysis Plot}\label{fig:basic}
\end{figure}

\bigskip



## Multilevel Bayesian Analysis

Now we move on to the multilevel analysis by taking the treatment assignment for each age group as our local-level prior information. Lupu (2013) argues that the less informed a person is, the more likely the effect of treatment will be. With the party platform treatment, younger voters are expected to identify more with a party than older adults do. We examine this prediction with Bayesian analysis under multilevel model (MLM). We treat each age group as the first-level, or local level, information and the entire population as the second-level information. The previous section also utilized Bayesian analysis but age group information was not handled thoroughly. This section overcomes the limitation of the previous section with MLM analysis. We first create a new column of age groups following criteria from the original article and divide the population into five groups: twenties, thirties, forties, fifties, and older. At the local level, we calculate the mean treatment proportion for each group as our local-level information (‘umat’) and then conduct MLM analysis with an explanatory variable of a global mean distance (‘frm2MdG’).  We use the global mean distance as our explanatory variable following our practice from the Quantitative Analaysis class. We also provide results from analyses with the original binary variable.

\begin{table}[]
\centering
\caption{Results from Bayesian MLM Analyses}
\label{table:mlm1}
\begin{tabular}{l|lllll} \hline
      & twenties & thirties & forties & fifties & older \\ \hline
 MLM with MdG & 0.0465 & 0.1704 &  -0.0068 & 0.2250 &  0.0707    \\
MLM with Binary IV & 0.0548  & 0.1661 & -0.0055 & 0.1595 & 0.0816  \\
MLM with Logistic Analysis & 0.2022 & 0.8020 &	0.1791 & 0.7290	&	0.2889 \\ \hline \hline
\end{tabular}
\end{table}

\bigskip


Table \ref{table:mlm1} shows results for each age group from Bayesian Analyses. The original article predicts that the effects of the first treatment will be largest with younger voters and smallest with the oldest group. In contrast, results from MLM show a different pattern and require a more complexed interpretation. First, the coefficients across different age groups are not linear: treatment effects fluctuate as we move from younger to older generations. Second, the direction of effects contradicts the prediction: the effects are largest with the fifties and smallest with forties. The effects on youngest group of twenties are placed as the third highest, locating the group in the middle among the five groups. Other MLM analyses show similar results and patterns. The second row of Table \ref{table:mlm1} reports results with a binary explanatory variable. The only different can be found between the thirties and fifties, the former replacing the latter as the highest treatment effects. 

We conduct a final Bayesian analysis with a Bernoulli-binomial distribution model. While the original article presents results with differences in means and does not provide additional analyses with logistic regression, the binary outcome variable makes it possible to conduct logistic analysis as well. The last row of Table \ref{table:mlm1} presents these results. The degree of treatment effects matches the pattern of those in the second row: treatment effects are largest among thirties and smallest among forties; and the coefficients do not correspond to the prediction of the original article. We provide additional analyses on results with tables and plots of posterior distributions of MLM with normal and logistic regressions. 


\begin{figure}[H]
\centering
\includegraphics[width=13cm, height=12cm]{normal mlm.pdf}
\caption{MLM Normal Bayesian Analysis Plot}\label{fig:normal}
\end{figure}

\bigskip


\begin{figure}[H]
\centering
\includegraphics[width=13cm, height=12cm]{logistic mlm.pdf}
\caption{MLM Logistic Bayesian Analysis Plot}\label{fig:logistic}
\end{figure}

\bigskip

The plots illustrate distribution of coefficients from Bayesian analysis. Figure \ref{fig:normal} shows results from the multilevel normal model. The posterior distributions of each group are centered around the mean but it does not follow the normal distribution. Coefficients are more clustered around the mean than a normal distribution except for the fifties group, where a bimodal distribution can be found. Figure \ref{fig:logistic} illustrates posterior distributions from the logistic model. Results from this model show starker similarity to the bimodal distribution than those with the normal model. 


\begin{table}[]
\centering
\caption{MLM Normal Bayesian Analysis Results}
\label{table:normalmlm}
\begin{tabular}{llllllll} \hline
              & mean    & se\_mean & sd     & 2.50\%  & 97.50\% & n\_eff & Rhat   \\ \hline
beta{[}1,1{]} & 0.3114  & 0.0206   & 0.0532 & 0.1916  & 0.3878  & 7      & 1.1798 \\
beta{[}1,2{]} & 0.1595  & 0.0356   & 0.0813 & 0.0549  & 0.3321  & 5      & 1.2605 \\
beta{[}2,1{]} & 0.3515  & 0.0172   & 0.0511 & 0.2359  & 0.4327  & 9      & 1.1266 \\
beta{[}2,2{]} & -0.0055 & 0.0076   & 0.0689 & -0.1314 & 0.1431  & 82     & 1.0415 \\
beta{[}3,1{]} & 0.3953  & 0.0021   & 0.0562 & 0.2747  & 0.5158  & 692    & 1.0112 \\
beta{[}3,2{]} & 0.0816  & 0.0033   & 0.0804 & -0.0921 & 0.2468  & 588    & 1.0125 \\
beta{[}4,1{]} & 0.376   & 0.0277   & 0.0656 & 0.2449  & 0.4754  & 6      & 1.2164 \\
beta{[}4,2{]} & 0.1661  & 0.0037   & 0.0807 & 0.0044  & 0.3435  & 483    & 1.0235 \\
beta{[}5,1{]} & 0.4153  & 0.0176   & 0.0519 & 0.3112  & 0.5075  & 9      & 1.1253 \\
beta{[}5,2{]} & 0.0548  & 0.0163   & 0.072  & -0.0865 & 0.2083  & 20     & 1.0629 \\ \hline
\end{tabular}
\end{table}

\bigskip

\begin{table}[]
\centering
\caption{MLM Stan Results for Logistic Regression}
\label{table:logitmlm}
\begin{tabular}{llllllll} \hline
              & mean     & se\_mean & sd      & 2.50\%   & 97.50\%  & n\_eff & Rhat    \\ \hline
beta{[}1,1{]} & -0.8414  & 0.07933  & 0.23706 & -1.35136 & -0.47404 & 9      & 1.37097 \\
beta{[}1,2{]} & 0.72904  & 0.07357  & 0.30906 & 0.18525  & 1.30185  & 18     & 1.35605 \\
beta{[}2,1{]} & -0.74078 & 0.02589  & 0.19007 & -1.09755 & -0.34119 & 54     & 1.06943 \\
beta{[}2,2{]} & 0.17907  & 0.08907  & 0.27162 & -0.38941 & 0.74443  & 9      & 1.34686 \\
beta{[}3,1{]} & -0.41877 & 0.09478  & 0.23508 & -0.93312 & -0.002   & 6      & 1.27272 \\
beta{[}3,2{]} & 0.28886  & 0.12016  & 0.30762 & -0.26348 & 0.91516  & 7      & 1.21744 \\
beta{[}4,1{]} & -0.62649 & 0.14647  & 0.27934 & -1.1373  & -0.07156 & 4      & 1.50022 \\
beta{[}4,2{]} & 0.80207  & 0.25671  & 0.43166 & 0.03296  & 1.44284  & 3      & 1.87041 \\
beta{[}5,1{]} & -0.4097  & 0.1147   & 0.23604 & -0.81481 & 0.03231  & 4      & 1.41331 \\
beta{[}5,2{]} & 0.20222  & 0.13027  & 0.30375 & -0.45923 & 0.83473  & 5      & 1.32433 \\ \hline
\end{tabular}
\end{table}


We approach to the disparity in posterior distributions with Rstan results tables. Tables \ref{table:normalmlm} and \ref{table:logitmlm} show mean, standard deviation, credible intervals, effective number of samples, and Rhat from MCMC processes. Rhat contains information of convergence and mix of four chains of a Bayesian analysis, showing how all the four chains return corresponding distribution of coefficients. If Rhat is close to 1, we refer those chains converge and mix properly. Rhat values from Table \ref{table:normalmlm} are close to 1 but the fifties group, reflecting non-normal distribution in Figure \ref{fig:normal}. Rhat values in Table \ref{table:logitmlm} show that they become even larger with the logistic MLM analysis. An additional feature from these approaches is the effective sample size and they are drastically smaller with the logistic analysis. Questions remain with these results regarding the cause of these problems: the feasibility of conducting Bayesian analysis with binary variables, compatibility of logistic and normal distribution in the inference, and the interpretation of bimodality in posterior distribution. 


# Conclusion
This paper suggests a preliminary proposal of statistical inference to the 'Party Brand' article [@Lupu2013]. While the original article constructs a model based on Bayesian psychological theory, the article is limited in practicing actual Bayesian statistical analysis. We overcome this limitation by providing results from the actual analysis and find contradicting results from the original article. Contrary to the prediction from the psychological model of low information prior, we find that age groups are not the determining factor of treatment effects. Rather, the pattern of treatment effects fluctuates as we move from younger to older age groups. This paper applies the Bayesian statistical analysis based on the understanding of potential outcomes and of filling in the missing observations. While the detailed processes we adopted may entail flaws and errors, we expect that our analysis will provide us with a better understanding on causal inferences from statistical analyses. 


# Code Appendix
The Rmarkdown files for the text and the statistical analysis, and the R data file can be accessed at the author's github repository: https://github.com/skim7713/PS531-Final.

# Reference
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

